---
title: "assignment1_R_olympic"
output: html_notebook
---
```{r}
library(dplyr)
data <- read.csv('athlete_events.csv')
head(data)
```

1. How old were the youngest male and female participants of the 1996 Olympics?
- 16 and 15
- 14 and 12
- 16 and 12
- 13 and 11
```{r}
youngest_male <- data[data['Sex'] == 'M' & data['Year'] == '1996', 'Age']
m <- which.min(youngest_male)
youngest_male[m]
```
```{r}
youngest_female <- data[data['Sex'] == 'F' & data['Year'] == '1996', 'Age']
f <- which.min(youngest_female)
youngest_female[f]
```

2. What was the percentage of male gymnasts among all the male participants of the 2000 Olympics? Round the answer to the first decimal.
Hint: here and further if needed drop duplicated sportsmen to count only unique ones.
- 0.2
- 1.5
- 2.5
- 7.7
```{r}
target <- data[data['Sex'] == 'M' & data['Sport'] == 'Gymnastics' & data['Year'] == 2000, 'ID']
all <- data[data['Sex'] == 'M' & data['Year'] == 2000, 'ID']
percentage <- length(target[!duplicated(target[])])/length(all[!duplicated(all[])])
round(percentage*100, 1)
```

3. What are the mean and standard deviation of height for female basketball players participated in the 2000 Olympics? Round the answer to the first decimal.
- 178.5 and 7.2
- 179.4 and 10
- 180.7 and 6.7
- 182.4 and 9.1
```{r}
m <- data[data['Sex'] == 'F' & data['Sport'] == 'Basketball' & data['Year'] == 2000, 'Height']
round(mean(m), 1)
```
```{r}
s <- data[data['Sex'] == 'F' & data['Sport'] == 'Basketball' & data['Year'] == 2000, 'Height']
round(sd(s), 1)
```

4. Find a sportsperson participated in the 2002 Olympics, with the highest weight among other participants of the same Olympics. What sport did he or she do?
- Judo
- Bobsleigh
- Weightlifting
- Boxing
```{r}
player <- data[data['Year'] == 2002, 'Weight']
m <- which.max(player)
player <- player[m]
d <- data[data['Weight'] == player & data['Year'] == 2002, 'Sport']
d[!is.na(d)]
```

5. How many times did Pawe Abratkiewicz participate in the Olympics held in different years?
- 0
- 1
- 2
- 3
```{r}
target <- data[data['Name'] == 'Pawe Abratkiewicz', 'Year']
length(target[!duplicated(target[])])
```

6. How many silver medals in tennis did Australia win at the 2000 Olympics?
- 0
- 1
- 2
- 3
```{r}
target <- data[data['Medal'] == 'Silver' & data['Sport'] == 'Tennis' & data['Team'] == 'Australia' & data['Year'] == 2000, 'ID']
target <- target[!is.na(target)]
length(target)
```

7. Is it true that Switzerland won fewer medals than Serbia at the 2016 Olympics? Do not consider NaN values in Medal column.
- Yes
- No
```{r}
switzerland <- data[data['Team'] == 'Switzerland' & data['Year'] == 2016, 'Medal']
switzerland <- switzerland[!is.na(switzerland)]
switzerland <- length(switzerland)
serbia <- data[data['Team'] == 'Serbia' & data['Year'] == 2016, 'Medal']
serbia <- serbia[!is.na(serbia)]
serbia <- length(serbia)
if (switzerland < serbia) {
    print('Yes')
} else {
    print('No')
}
```

8. What age category did the fewest and the most participants of the 2014 Olympics belong to?
- [45-55] and [25-35) correspondingly
- [45-55] and [15-25) correspondingly
- [35-45] and [25-35) correspondingly
- [45-55] and [35-45) correspondingly
```{r}
d <- data[data['Year'] == 2014, ]
d <- d %>% select(1, 4)
d <- unique(d)
d <- d$Age
d <- cut(d, breaks = 4,right=FALSE)
table(d)
```

9. Is it true that there were Summer Olympics held in Lake Placid? Is it true that there were Winter Olympics held in Sankt Moritz?
- Yes, Yes
- Yes, No
- No, Yes
- No, No
```{r}
target <- data[data['Season'] == 'Summer' & data['City'] == 'Lake Placid', 'ID']
target <- length(target)
if (target == 0) {
    print('No')
} else {
    print('Yes')
}
```
```{r}
target <- data[data['Season'] == 'Winter' & data['City'] == 'Sankt Moritz', 'ID']
target <- length(target)
if (target == 0) {
    print('No')
} else {
    print('Yes')
}
```

10. What is the absolute difference between the number of unique sports at the 1995 Olympics and 2016 Olympics?
- 16
- 24
- 26
- 34
```{r}
first <- data[data['Year'] == 1995, 'Sport']
first <- first[!is.na(first)]
first <- length(first[!duplicated(first[])])
second <- data[data['Year'] == 2016, 'Sport']
second <- second[!is.na(second)]
second <- length(second[!duplicated(second[])])
dif <- abs(first - second)
dif
```
