---
title: "assignment2_USA_flights"
output: html_notebook
---
```{r}
library(dplyr)
library(plotly)
```
```{r}
flights_df <- read.csv('2008.csv')
head(flights_df)
```

1. Find top-10 carriers in terms of the number of completed flights (UniqueCarrier column)?
Which of the listed below is not in your top-10 list?
- DL
- AA
- OO
- EV
```{r}
d <- flights_df$UniqueCarrier
tbl <- sort(table(d), decreasing = TRUE) 
tbl[1:10]
```

2. Plot distributions of flight cancellation reasons (_CancellationCode_).
What is the most frequent reason for flight cancellation? (Use this link to translate codes into reasons)
- carrier
- weather conditions
- National Air System
- security reasons
```{r}
t <- table(flights_df$CancellationCode)
t <- t[2:5]
barplot(t, col = 'orange')
```

3. Which route is the most frequent, in terms of the number of flights?
(Take a look at 'Origin' and 'Dest' features. Consider A->B and B->A directions as different routes)
- New-York – Washington
- San-Francisco – Los-Angeles
- San-Jose – Dallas
- New-York – San-Francisco
```{r}
d <- flights_df
d$Flight <- paste(d$Origin, d$Dest, sep="-")
sort(table(d$Flight), decreasing = TRUE)[1]
```

4. Find top-5 delayed routes (count how many times they were delayed on departure). From all flights on these 5 routes, count all flights with weather conditions contributing to a delay.
- 449
- 539
- 549
- 668
```{r}
d <- flights_df
top <- d[!is.na(d$DepDelay) & d$DepDelay > 0, c('Origin', 'Dest', 'DepDelay')]
top <- top %>% group_by(Origin, Dest) %>% summarise(count = n()) %>% arrange(desc(count))
top <- top[c(1:5),]
top
```
```{r}
first <- d[d['Origin'] == 'LAX' & d['Dest'] == 'SFO' &  !(is.na(d$WeatherDelay)) & d$WeatherDelay > 0, ]
second <- d[d['Origin'] == 'DAL' & d['Dest'] == 'HOU' &  !(is.na(d$WeatherDelay)) & d$WeatherDelay > 0, ]
third <- d[d['Origin'] == 'SFO' & d['Dest'] == 'LAX' &  !(is.na(d$WeatherDelay)) & d$WeatherDelay > 0, ]
fourth <- d[d['Origin'] == 'ORD' & d['Dest'] == 'LGA' &  !(is.na(d$WeatherDelay)) & d$WeatherDelay > 0, ]
fifth <- d[d['Origin'] == 'HOU' & d['Dest'] == 'DAL' &  !(is.na(d$WeatherDelay)) & d$WeatherDelay > 0, ]
s <- count(first) + count(second) + count(third) + count(fourth) + count(fifth)
s
```

5. Examine the hourly distribution of departure times. For that, create a new series from DepTime, removing missing values.
Choose all correct statements:
- Flights are normally distributed within time interval [0-23] (Search for: Normal distribution, bell curve).
- Flights are uniformly distributed within time interval [0-23].
- In the period from 0 am to 4 am there are considerably less flights than from 7 pm to 8 pm.
```{r}
d <- flights_df
d <- round(d['DepTime']/100)
d <- unlist(d)
hist(d, col="blueviolet", breaks=24)
```

6. Show how the number of flights changes through time (on the daily/weekly/monthly basis) and interpret the findings.
Choose all correct statements:
- The number of flights during weekends is less than during weekdays (working days).
- The lowest number of flights is on Sunday.
- There are less flights during winter than during summer.
```{r}
d <- flights_df
d <- table(d['DayOfWeek'])
barplot(d, col='cyan2')
```
```{r}
d <- flights_df
d <- table(d['Month'])
barplot(d, col='aquamarine2')
```

7. Examine the distribution of cancellation reasons with time. Make a bar plot of cancellation reasons aggregated by months.
Choose all correct statements:
- December has the highest rate of cancellations due to weather.
- The highest rate of cancellations in September is due to Security reasons.
- April's top cancellation reason is carriers.
- Flights cancellations due to National Air System are more frequent than those due to carriers.
```{r}
d <- flights_df
d <- d[d$Cancelled != 0, c('CancellationCode','Month')]
d <- d %>% group_by(CancellationCode, Month) %>% count(CancellationCode)
c_reasons <- d[d$CancellationCode == 'A',]
barchart <- plot_ly(x = c_reasons$Month, y = c_reasons$n, type = "bar")
c_reasons <- d[d$CancellationCode == 'B',]
barchart <- barchart %>% add_trace(y = c_reasons$n)
c_reasons <- d[d$CancellationCode == 'C',]
barchart <- barchart %>% add_trace(y = c_reasons$n)
c_reasons <- vector('integer', 12)
months <- c(d[d$CancellationCode == 'D',]['Month'])
for (i in (1:12)) {
    if (i %in% months$Month) {
        c_reasons[i] = d[d$CancellationCode == 'D' & d$Month == i,]['n']
    }
}
barchart <- barchart %>% add_trace(y = c_reasons)
barchart
```

8. Which month has the greatest number of cancellations due to Carrier?
- May
- January
- September
- April
```{r}
d <- flights_df
d <- d[d$Cancelled == 1 & d$CancellationCode == 'A', c('CancellationCode','Month', 'UniqueCarrier')]
d <- d %>% group_by(Month) %>% summarise(count = n())
barplot(d$count, col='deeppink4')
```

9. Identify the carrier with the greatest number of cancellations due to carrier in the corresponding month from the previous question.
- 9E
- EV
- HA
- AA
```{r}
d <- flights_df
d <- d[d['Month'] == 4 & d['Cancelled'] == 1, ]
d <- d$UniqueCarrier
tbl <- sort(table(d), decreasing = TRUE) 
tbl[1]
```

10. Examine median arrival and departure delays (in time) by carrier. Which carrier has the lowest median delay time for both arrivals and departures? Leave only non-negative values of delay times ('ArrDelay', 'DepDelay'). Boxplots can be helpful in this exercise, as well as it might be a good idea to remove outliers in order to build nice graphs. You can exclude delay time values higher than a corresponding .95 percentile.
- EV
- OO
- AA
- AQ
```{r}
d <- flights_df
d <- d[!is.na(d$ArrDelay) & 0 <= d$ArrDelay & d$ArrDelay <= quantile(d$ArrDelay, 0.9, na.rm = TRUE), c('ArrDelay', 'UniqueCarrier')]
d <- d %>% group_by(UniqueCarrier) %>% table
carriers <- names(temp[1,])
boxplot <- plot_ly(type = "box")
for (i in (1:length(carriers))) {
    boxplot <- boxplot %>% add_boxplot(y = d[, carriers[i]], name=carriers[i])
}
boxplot
```
```{r}
d <- flights_df
d <- d[!is.na(d$DepDelay) & 0 <= d$DepDelay & d$DepDelay <= quantile(d$DepDelay, 0.9, na.rm = TRUE), c('DepDelay', 'UniqueCarrier')]
d <- d %>% group_by(UniqueCarrier) %>% table
carriers <- names(temp[1,])
boxplot <- plot_ly(type = "box")
for (i in (1:length(carriers))) {
    boxplot <- boxplot %>% add_boxplot(y = d[, carriers[i]], name=carriers[i])
}
boxplot
```
